### 2025-01-28 manuscript/manuscript.qmd
updated manuscript terms

To capture key details of each study, we added additional fields to BibTeX entries for each study. Fields included information about the genre/type of stimuli employed, along with their duration and number; the number of analyzed features; and the model type, validation procedure and output measures. Additionally, we included study results using executable *R* code containing custom functions for meta-analysis. For complete details about our encoding procedure, see `studies/extraction_details.qmd` .

@@

|            | Regression                      | Classification            | Total |
|:-----------|:--------------------------------|:--------------------------|:------|
| Study N    | 22                              | 12                        | 34    |
| Model N    | 204                             | 86                        | 290   |
| Techniques | Neural Nets (NN): 66            | 25                        | 91    |
| Techniques | Flexible Discriminants (FD): 50 | 26                        | 76    |
| Techniques | Linear Methods (LM): 62         | 13                        | 75    |
| Techniques | Random Forests (RF): 14         | 10                        | 34    |
| Techniques | KS, Add. & KNN (KS): 12         | 12                        | 24    |
| Feature N  | Min=3, Md=605, Max=6670         | Min=8, Md=126, Max=8904   | NA    |
| Stimulus N | Min=20, Md=324, Max=2486        | Min=124, Md=300, Max=5192 | NA    |

Although the total number of studies meeting the criteria described in the previous section is modest (34 in total), they encompass a large array of models (290 in total) with a relatively even distribution among the three most popular techniques: flexible discriminants, neural nets, and linear methods. The number of features and stimuli within these studies varies significantly, ranging from as few as three features [@battcock2021in] to a maximum of almost 9000 features [@zhang2023mo]. The median number of features differs between regression (605) and classification (126) studies, primarily reflecting the nature of the datasets used in each approach. The number of stimuli is typically around 300-400 (with a median of 324 for regression and 300 for classification), though there is substantial variation, with the extremes from 20 stimuli in @beveridge2018po to 5192 stimuli in @alvarez2023ri. There are also additional dimensions to consider, such as the type of cross-validation used, the music genres analyzed (whether a single genre, multiple genres, or a mix), the type of journal in which the studies were published, and the extraction tool used to extract features. However, these variables do not lend themselves to a simple summary, so we will revisit them during the interpretation and discussion stages.

@@

| Concept      | Models, obs | $r$ \[95%-CI\]          | $t$   | $p$   | $\tau^2$ | $I^2$ |
|:-------------|:------------|:------------------------|:------|:------|:---------|:------|
| Valence All  | 102,60017   | 0.583 \[0.541-0.623\]   | 21.40 | .0001 | 0.094    | 97.7% |
| Valence Max  | 22,14172    | 0.669 \[0.560-0.755\]   | 9.58  | .0001 | 0.148    | 98.4% |
| *N Features* |             |                         |       |       |          |       |
| \<18 F       | 5,3036      | 0.811 \[0.542-0.823\]   | \-    | \-    | 0.182    | 98.9% |
| 18-260 F     | 11,7042     | 0.584 \[0.534-0.931\]   | \-    | \-    | 0.133    | 97.6% |
| 260+ F       | 5,3494      | 0.710 \[0.542-0.823\]   | \-    | \-    | 0.044    | 97.2% |
| *Techniques* |             |                         |       |       |          |       |
| KS           | 1,1838      | 0.466 \[-0.634-0.942\]  | \-    | \-    | 0.0185   | 95.1% |
| LM           | 8,1762      | 0.797 \[0.614-0.899\]   | \-    | \-    | 0.1370   | 96.4% |
| FD           | 6,4993      | 0.656 \[0.484-0.779\]   | \-    | \-     | 0.0574  | 96.9% |
| NN           | 4,2249      | 0.393 \[-0.355-0.835\]  | \-    | \-     | 0.0761  | 97.1% |
| RF           | 3,3330      | 0.750 \[0.292-0.928\]   | \-    | \-     | 0.093   | 98.5% |

<!-- table updated TE 21/1/2025 -->

@@

When analyzing the studies across the five modeling techniques used, the prediction do differ significantly (_Q_(4) = 45.7, _p_ < .0001). Notably, linear models (LM) and flexible discriminants (FD) were the most common, with 8 and 6 studies, respectively, allowing for more confident interpretations. Linear models achieved the highest prediction rate (_r_ = 0.784, 95% CI: 0.625–0.881), though this may be influenced by the smaller datasets typically used in these studies. These studies also exhibited substantially higher heterogeneity ($\tau^2$ = 0.137) compared to other techniques, where $\tau^2$ values were less than half of this. Flexible discriminants performed at a similar level to the overall model average (_r_ = 0.656, 95% CI: 0.484–0.779). While there were only 3 studies involving random forests, these performed well, achieving _r_ = 0.750, 95% CI: 0.292–0.928), and the relatively poor performance of the neural network models represented in four studies (_r_ = 0.340, 95% CI: -0.097–0.668) is difficult to explain without a deeper examination of the specific model architectures and the stimuli used in these studies. Kernel smoothing and KNN models (with only 1 study represented) achieved relatively poor success as well, _r_ = 0.383, 95% CI: 0.348–0.426).

@@

| Concept      | Models, obs | $r$ \[95%-CI\]            | $t$  | $p$    | $\tau^2$ | $I^2$ |
|:-------------|:------------|:--------------------------|:-----|:-------|:---------|:------|
| Arousal All  | 102, 60017  | 0.791 \[0.770-0.810\]     | 39.9 | 0.0001 | 0.069    | 96.2% |
| Arousal Max  | 22, 14172   | 0.809 \[0.740-0.860\]     | 13.6 | 0.0001 | 0.141    | 97.9% |
| *N Features* |             |                           |      |        |          |       |
| \<18 F       | 4, 3016     | 0.817 \[0.361-0.958\]     |      |        | 0.2295   | 99.2% |
| 18-260 F     | 10, 4222    | 0.586 \[0.356-0.748\]     |      |        | 0.1542   | 96.9% |
| 260+ F       | 7, 6190     | 0.702 \[0.612-0.774\]     |      |        | 0.0285   | 96.6% |
| *Techniques* |             |                           |      |        |          |       |
| KS           | 1, 1838     | 0.819 \[0.803-0.833\]     |      |        |  -       | -     |
| LM           | 8, 1762     | 0.881 \[0.808-0.928\]     |      |        | 0.0846   | 93.3% |
| FD           | 6, 4993     | 0.808 \[0.644-0.901\]     |      |        | 0.1125   | 98.0% |
| NN           | 4, 2249     | 0.533 \[0.328-0.691\]     |      |        | 0.0190   | 85.8% |
| RF           | 3, 3330     | 0.809 \[0.733-0.864\]     |      |        | 0.0025   | 65.4% |

>>>>

To capture key details of each study, we added additional fields to BibTeX entries for each study. Fields included information about the genre/type of stimuli employed, along with their duration and number; the number of analyzed features; and the model type -- Neural Nets (NN), Support Vector Machines (SVM), Linear Methods (LM), Tree-based Methods (TM), KS, Add. and KNN Methods (KM) -- validation procedure and output measures. Additionally, we included study results using executable *R* code containing custom functions for meta-analysis. For complete details about our encoding procedure, see `studies/extraction_details.qmd` .

@@

|            | Regression                      | Classification            | Total |
|:-----------|:--------------------------------|:--------------------------|:------|
| Study N    | 22                              | 12                        | 34    |
| Model N    | 204                             | 86                        | 290   |
| Techniques | Neural Nets: 66                 | 25                        | 91    |
| Techniques | Support Vector Machines: 50     | 26                        | 76    |
| Techniques | Linear Methods: 62              | 13                        | 75    |
| Techniques | Tree-based methods: 14          | 10                        | 34    |
| Techniques | KS, Add. & KNN: 12              | 12                        | 24    |
| Feature N  | Min=3, Md=605, Max=6670         | Min=8, Md=126, Max=8904   | NA    |
| Stimulus N | Min=20, Md=324, Max=2486        | Min=124, Md=300, Max=5192 | NA    |

Although the total number of studies meeting the criteria described in the previous section is modest (34 in total), they encompass a large array of models (290 in total) with a relatively even distribution among the three most popular techniques: Neural Nets, SVMs, and Linear Methods. The number of features and stimuli within these studies varies significantly, ranging from as few as three features [@battcock2021in] to a maximum of almost 9000 features [@zhang2023mo]. The median number of features differs between regression (605) and classification (126) studies, primarily reflecting the nature of the datasets used in each approach. The number of stimuli is typically around 300-400 (with a median of 324 for regression and 300 for classification), though there is substantial variation, with the extremes from 20 stimuli in @beveridge2018po to 5192 stimuli in @alvarez2023ri. There are also additional dimensions to consider, such as the type of cross-validation used, the music genres analyzed (whether a single genre, multiple genres, or a mix), the type of journal in which the studies were published, and the extraction tool used to extract features. However, these variables do not lend themselves to a simple summary, so we will revisit them during the interpretation and discussion stages.

@@

| Concept      | Models, obs | $r$ \[95%-CI\]          | $t$   | $p$   | $\tau^2$ | $I^2$ |
|:-------------|:------------|:------------------------|:------|:------|:---------|:------|
| Valence All  | 102,60017   | 0.583 \[0.541-0.623\]   | 21.40 | .0001 | 0.094    | 97.7% |
| Valence Max  | 22,14172    | 0.669 \[0.560-0.755\]   | 9.58  | .0001 | 0.148    | 98.4% |
| *N Features* |             |                         |       |       |          |       |
| \<18 F       | 5,3036      | 0.811 \[0.542-0.823\]   | \-    | \-    | 0.182    | 98.9% |
| 18-260 F     | 11,7042     | 0.584 \[0.534-0.931\]   | \-    | \-    | 0.133    | 97.6% |
| 260+ F       | 5,3494      | 0.710 \[0.542-0.823\]   | \-    | \-    | 0.044    | 97.2% |
| *Techniques* |             |                         |       |       |          |       |
| KS           | 1,1838      | 0.466 \[-0.634-0.942\]  | \-    | \-    | 0.0185   | 95.1% |
| LM           | 8,1762      | 0.797 \[0.614-0.899\]   | \-    | \-    | 0.1370   | 96.4% |
| SVM          | 6,4993      | 0.656 \[0.484-0.779\]   | \-    | \-     | 0.0574  | 96.9% |
| NN           | 4,2249      | 0.393 \[-0.355-0.835\]  | \-    | \-     | 0.0761  | 97.1% |
| TM           | 3,3330      | 0.750 \[0.292-0.928\]   | \-    | \-     | 0.093   | 98.5% |

<!-- table updated TE 21/1/2025, revised terms 28/1/2025 -->

@@

When analyzing the studies across the five modeling techniques used, the prediction do differ significantly (_Q_(4) = 45.7, _p_ < .0001). Notably, linear models (LM) and Support Vector Machines (SVM) were the most common, with 8 and 6 studies, respectively, allowing for more confident interpretations. Linear models achieved the highest prediction rate (_r_ = 0.784, 95% CI: 0.625–0.881), though this may be influenced by the smaller datasets typically used in these studies. These studies also exhibited substantially higher heterogeneity ($\tau^2$ = 0.137) compared to other techniques, where $\tau^2$ values were less than half of this. Flexible discriminants performed at a similar level to the overall model average (_r_ = 0.656, 95% CI: 0.484–0.779). While there were only 3 studies involving random forests, these performed well, achieving _r_ = 0.750, 95% CI: 0.292–0.928), and the relatively poor performance of the neural network models represented in four studies (_r_ = 0.340, 95% CI: -0.097–0.668) is difficult to explain without a deeper examination of the specific model architectures and the stimuli used in these studies. Kernel smoothing and KNN models (with only 1 study represented) achieved relatively poor success as well, _r_ = 0.383, 95% CI: 0.348–0.426).

@@

| Concept      | Models, obs | $r$ \[95%-CI\]            | $t$  | $p$    | $\tau^2$ | $I^2$ |
|:-------------|:------------|:--------------------------|:-----|:-------|:---------|:------|
| Arousal All  | 102, 60017  | 0.791 \[0.770-0.810\]     | 39.9 | 0.0001 | 0.069    | 96.2% |
| Arousal Max  | 22, 14172   | 0.809 \[0.740-0.860\]     | 13.6 | 0.0001 | 0.141    | 97.9% |
| *N Features* |             |                           |      |        |          |       |
| \<18 F       | 4, 3016     | 0.817 \[0.361-0.958\]     |      |        | 0.2295   | 99.2% |
| 18-260 F     | 10, 4222    | 0.586 \[0.356-0.748\]     |      |        | 0.1542   | 96.9% |
| 260+ F       | 7, 6190     | 0.702 \[0.612-0.774\]     |      |        | 0.0285   | 96.6% |
| *Techniques* |             |                           |      |        |          |       |
| KS           | 1, 1838     | 0.819 \[0.803-0.833\]     |      |        |  -       | -     |
| LM           | 8, 1762     | 0.881 \[0.808-0.928\]     |      |        | 0.0846   | 93.3% |
| SVM          | 6, 4993     | 0.808 \[0.644-0.901\]     |      |        | 0.1125   | 98.0% |
| NN           | 4, 2249     | 0.533 \[0.328-0.691\]     |      |        | 0.0190   | 85.8% |
| TM           | 3, 3330     | 0.809 \[0.733-0.864\]     |      |        | 0.0025   | 65.4% |
